---
title: "Unsupervised Learning"
output: html_notebook
---

# Prerequisites

```{r slide-5}
## ----package-requirements------------------------------------------------
library(tidyverse)   # data munging & visualization
library(cluster)     # additional clustering techniques
library(factoextra)  # clustering & PCA visualizations

## ----data-requirements---------------------------------------------------
USArrests                 # primary example data
AmesHousing::make_ames()  # a few additional examples
```

# Clustering

## Prepare our data for _k_-means

1. Rows are observations (individuals) and columns are variables.
2. Any missing value in the data must be removed or estimated.
3. The data must be standardized (centered at mean zero and scaled to one standard deviation) to make variables comparable.

```{r slide-15}
crime <- USArrests %>%
  drop_na() %>%
  scale() %>%
  print()
```

## Applying k-means

```{r slide-16}
k3 <- kmeans(crime, centers = 3, nstart = 20)

# tidied output
broom::tidy(k3)

# full model output
glimpse(k3)
```


## Interpreting output 

```{r slide-17}
as_tibble(crime) %>%
  mutate(
    cluster = k3$cluster, 
    label = paste0(row.names(USArrests), " (", cluster, ")")
    ) %>% 
  gather(Crime, Rate, Murder, Assault, Rape) %>%
  ggplot(aes(UrbanPop, Rate, color = factor(cluster), label = label)) +
  geom_text(show.legend = FALSE) +
  facet_wrap(~ Crime) +
  ylab("arrests per 100,000 residents (standardized)")
```


## Determining optimal clusters

Fill in the following code chunk to identify the optimal number of clusters for our crime data.

```{r slide-23a}
# use the elbow heuristic (hint: "wss")
fviz_nbclust(____, kmeans, method = "___", k.max = 20)
```

```{r slide-23b}
# use the gab stat (hint: "gap_stat")
fviz_nbclust(____, kmeans, method = "___", k.max = 20, verbose = FALSE)
```


## Applying hierarchical clustering


```{r slide-33a}
# you can use hclust
option1 <- hclust(dist(crime), method = "average")
plot(option1)
```

```{r slide-33b}
# or agnes
option2 <- cluster::agnes(crime, method = "average")
plot(option2, which = 2)
```

## Which method to use?

The `agnes()` function provides an ___agglomerative coefficent___, which measures the amount of clustering structure found (values closer to 1 suggest strong clustering structure).

```{r slide-35a}
# agglomerative coefficent from earlier agnes() model
option2$ac
```

This allows us to find certain hierarchical clustering methods that can identify stronger clustering structures.

```{r slide-35b}
# methods to assess
m <- c("average", "single", "complete", "ward")
names(m) <- m

# function to compute coefficient
ac <- function(x) {
  cluster::agnes(crime, method = x)$ac
}

# get agglomerative coefficient for each linkage method
map_dbl(m, ac)
```
]


## Which clusters to use?

Visual assessment

```{r slide-36a}
# re-run hc with Ward method
hc_ward <- cluster::agnes(crime, method = "ward")

# highlight 4 clusters
plot(hc_ward, which = 2)
rect.hclust(hc_ward, k = 4, border = 2:5)
```

Elbow method or Gap stat

```{r slide-36b}
# fviz_nbclust can use differen FUNctions
# see ?fviz_nbclust & ?hcut
fviz_nbclust(crime, FUN = hcut, method = "gap_stat", verbose = FALSE)
```


## Extract results and visualize

```{r slide-37}
as_tibble(crime) %>%
  mutate(
    cluster = cutree(hc_ward, k = 3), # use cutree to get clusters #<<
    label = paste0(row.names(USArrests), " (", cluster, ")")
    ) %>% 
  gather(Crime, Rate, Murder, Assault, Rape) %>%
  ggplot(aes(UrbanPop, Rate, color = factor(cluster), label = label)) +
  geom_text(show.legend = FALSE) +
  facet_wrap(~ Crime) +
  ylab("arrests per 100,000 residents (standardized)")
```


## Ohio's cluster

```{r slide-38}
# get ohio neighbors with k-means
df_kmeans <- as_tibble(crime) %>%
  mutate(
    cluster = kmeans(crime, 3, nstart = 20)$cluster, # use cutree to get clusters #<<
    label = paste0(row.names(USArrests))
    )

ohio_cluster <- filter(df_kmeans, label == "Ohio") %>% pull(cluster)

kmean_friends <- df_kmeans %>%
  filter(cluster == ohio_cluster) %>%
  arrange() %>%
  pull(label)

# get ohio neighbors with hierarchical
df_hc <- as_tibble(crime) %>%
  mutate(
    cluster = cutree(hc_ward, k = 3), # use cutree to get clusters #<<
    label = paste0(row.names(USArrests))
    )

ohio_cluster <- filter(df_hc, label == "Ohio") %>% pull(cluster)

hcut_friends <- df_hc %>%
  filter(cluster == ohio_cluster) %>%
  arrange() %>%
  pull(label)

# Common states between _k_-means and Hierarchical clustering
intersect(hcut_friends, kmean_friends)

# Different states between _k_-means and Hierarchical clustering
setdiff(union(hcut_friends, kmean_friends), intersect(hcut_friends, kmean_friends))
```


# Dimension Reduction via PCA

## Prepare our data for PCA

1. Rows are observations (individuals) and columns are variables.
2. Any missing value in the data must be removed or estimated.
3. The data must be standardized (centered at mean zero and scaled to one standard deviation) to make variables comparable.

```{r slide-47}
# exact same procedure you ran for clustering
crime <- USArrests %>%
  drop_na() %>%
  scale() %>%
  print()
```

4. Apply `prcomp()`

```{r slide-48}
# perform PCA
pca_result <- prcomp(crime, scale = FALSE)

# PCA model output
names(pca_result)
```

## Understanding PCs 

* `rotation` provides the principal component loadings
* There will be the same number of PCs as variables
* Positive vs negative direction:
   - By default, loadings (aka eigenvectors) in R point in the negative direction. 
   - Positive pointing eigenvectors are more intuitive 
   - To use the positive-pointing vector, we multiply the default loadings by -1.

```{r slide-49}
# convert loadings to positive
pca_result$rotation <- -pca_result$rotation

# there will be the same number of PCs as variables
pca_result$rotation
```

___Loadings represent coefficients; illustrates each variables influence on the principal component___

## Understanding PCs 

We can visualize these contributions

PC1 appears to represent violent crime

```{r slide-50a}
fviz_contrib(pca_result, choice = "var", axes = 1)
```

PC2 appears to represent urban density

```{r slide-50b}
fviz_contrib(pca_result, choice = "var", axes = 2)
```


## Understanding PCs 

* `x` provides the principal component scores
* The principal components scores simply places a standardized score for each observation for each principal component.
* Interpretation example: Alaska
   - PC1: 1.93 standard deviations above average value for PC1 (high violent crime)
   - PC2: 1 standard deviation below average value for PC2 (low urbanization)

```{r slide-51}
pca_result$x <- -pca_result$x
pca_result$x
```

## Selecting the number of PCs

* The goal of PCA is dimension reduction but how many PCs do we keep?

* Two common approaches:
   1. scree plot  
      a. Eigenvalue: retain PCs $\geq$ 1 (an eigenvalue of 1 means that the principal component would explain about one variable’s worth of the variability)  
      b. Variance explained: look for elbow

```{r slide-53}
fviz_screeplot(pca_result, choice = "eigenvalue")
fviz_screeplot(pca_result, choice = "variance")
```

   2. proportion of variance explained

```{r slide-54}
# compute eigenvalues
eigen <- pca_result$sdev^2

# compute the PVE of each principal component
PVE <- eigen / sum(eigen)

# how many PCs required to explain at least 90% of total variability
min(which(cumsum(PVE) >= .90))
```

___PVE provides us a technical way to identify the optimal number of principal components to keep based on the total variability that we would like to account for. In feature engineering, it is common for the default to be set at 95%.___

## Insight interpretation 


```{r slide-55a}
fviz_pca_var(pca_result, alpha.var = "contrib")
```

```{r slide-55b}
fviz_pca(pca_result, alpha.ind = .5, labelsize = 3, repel = TRUE)
```


## PCA with mixed data 

* Rarely do we have only numeric data
* But what do we do with categorical features?
* The answer is...it depends
* PCA for inference
   - option 1: one hot encode
   - option 2: use GLRMs 

One-hot encoding example:

```{r slide-57}
# full ames data set --> recode ordinal variables to numeric
ames_full <- AmesHousing::make_ames() %>%
  mutate_if(str_detect(names(.), "Qual|Cond|QC|Qu"), as.numeric)

# one-hot encode --> retain only the features and not sale price
full_rank  <- caret::dummyVars(Sale_Price ~ ., data = ames_full, fullRank = TRUE)
ames_1hot <- predict(full_rank, ames_full)

# new dimensions
dim(ames_1hot)

# apply PCA to one-hot encoded data
pca_one_hot <- prcomp(ames_1hot, scale = TRUE)

# sign adjustment to loadings and scores
pca_one_hot$rotation <- -pca_one_hot$rotation
pca_one_hot$x <- -pca_one_hot$x

# scree plot
fviz_screeplot(pca_one_hot, ncp = 20, choice = "eigenvalue")
```

* PCA for downstream modeling

```{r slide-58}
# get feature set
ames_full <- AmesHousing::make_ames()
features <- subset(ames_full, select = -Sale_Price)

# preprocess data
preprocess <- caret::preProcess(
  x = features,
  method = c("center", "scale", "pca"),
  thresh = 0.95
)

preprocess
```
