---
title: "Supervised Modeling Process"
output: html_notebook
---

# Prerequisites

```{r slide-4}
# Packages required
library(rsample)
library(caret)
library(tidyverse)

# Data required
## ames data
ames <- AmesHousing::make_ames()

## attrition data
churn <- rsample::attrition
```


# Mechanics of data splitting 

Two most common ways of splitting data include:

* simple random sampling: randomly select observations
* stratified sampling: preserving distributions
   - classification: sampling within the classes to preserve the 
     distribution of the outcome in the training and test sets
   - regression: determine the quartiles of the data set and sample within those
      artificial groups

```{r slide-8}
set.seed(123) # for reproducibility
split <- initial_split(diamonds, strata = "price", prop = 0.7)
train <- training(split)
test  <- testing(split)

# Do the distributions line up? 
ggplot(train, aes(x = price)) + 
  geom_line(stat = "density", 
            trim = TRUE) + 
  geom_line(data = test, 
            stat = "density", 
            trim = TRUE, col = "red")
```


# Your Turn! 

1. Use __rsample__ to split the Ames housing data (`ames`) and the Employee attrition data (`churn`) using stratified sampling and with a 80% split.
2. Verify that the distribution between training and test sets are similar.

```{r slide-9}
# ames data
set.seed(123)
ames_split <- initial_split(ames, prop = _____, strata = "Sale_Price")
ames_train <- training(_____)
ames_test  <- testing(_____)

# attrition data
set.seed(123)
churn_split <- initial_split(churn, prop = _____, strata = "Attrition")
churn_train <- training(_____)
churn_test  <- testing(_____)
```


# Putting the process together 

Let's put these pieces together and analyze the Ames housing data:

1. Split into training vs testing data
2. Specify a resampling procedure
3. Create our hyperparameter grid
4. Execute grid search
5. Evaluate performance

___This grid search takes ~2 min___

```{r slide-35}
# 1. stratified sampling with the rsample package
set.seed(123)
split  <- initial_split(ames, prop = 0.7, strata = "Sale_Price")
ames_train  <- training(split)
ames_test   <- testing(split)

# 2. create a resampling method
cv <- trainControl(
  method = "repeatedcv", 
  number = 10, 
  repeats = 5
  )

# 3. create a hyperparameter grid search
hyper_grid <- expand.grid(k = seq(2, 26, by = 2))

# 4. execute grid search with knn model
#    use RMSE as preferred metric
knn_fit <- train(
  Sale_Price ~ ., 
  data = ames_train, 
  method = "knn", 
  trControl = cv, 
  tuneGrid = hyper_grid,
  metric = "RMSE"
  )

# 5. evaluate results
# print model results
knn_fit

# plot cross validation results
ggplot(knn_fit$results, aes(k, RMSE)) + 
  geom_line() +
  geom_point() +
  scale_y_continuous(labels = scales::dollar)
```
