---
title: "Gradient Boosting Machines"
output: html_notebook
---

# Prereqs 

```{r slide-3}
library(gbm)
library(xgboost)
library(vip)
library(pdp)

# ames data
ames <- AmesHousing::make_ames()

# split data
set.seed(123)
split <- rsample::initial_split(ames, strata = "Sale_Price")
ames_train <- rsample::training(split)
```

# Basic GBM 

gbm package:

- The original R implementation of GMBs (by Greg Ridgeway)
- Slower than modern implementations (but still pretty fast)
- Provides OOB error estimate
- Supports the weighted tree traversal method for fast construction of PDPs

Let's run your first GBM model

```{r slide-17}
set.seed(123)
ames_gbm <- gbm(
  formula = Sale_Price ~ .,
  data = ames_train,
  distribution = "gaussian", # or bernoulli, multinomial, etc. 
  n.trees = 5000, 
  shrinkage = 0.1, 
  interaction.depth = 1, 
  n.minobsinnode = 10, 
  cv.folds = 5 
  )  

# find index for n trees with minimum CV error
min_MSE <- which.min(ames_gbm$cv.error)

# get MSE and compute RMSE
sqrt(ames_gbm$cv.error[min_MSE])
```

# Number of trees

- The averaging in bagging and RF makes it very difficult to overfit with too many trees
- GBMs will chase residuals as long as you allow them to
- Consequently:
   - We must provide enough trees to minimize error
   - But not too many where we begin to overfit
   - .red[plus, number of trees is dependent on other hyperparameters]

Use CV or OOB

```{r slide-22}
gbm.perf(ames_gbm, method = "cv") # or "OOB"
```

# Tuning strategy 

1. fix tree hyperparameters
    - moderate tree depth
    - default min obs
2. set our learning rate at .01
3. increase CV to ensure unbiased error estimate
4. Results
   - Lowest error rate yet ($21,914.55)!
   - Used nearly all our trees --> increase to 6000?
   - took ~ 2.25 min
5. Compared to learning rate of .001
   - error rate of $24,791.66
   - took ~ 4 min

This model run takes ~2 mins 

```{r slide-28}
set.seed(123)
ames_gbm1 <- gbm(
  formula = Sale_Price ~ .,
  data = ames_train,
  distribution = "gaussian", # or bernoulli, multinomial, etc. #<<
  n.trees = 5000, #<<
  shrinkage = 0.01, #<<
  interaction.depth = 3, #<<
  n.minobsinnode = 10, #<<
  cv.folds = 10 #<<
  )

# find index for n trees with minimum CV error
min_MSE <- which.min(ames_gbm1$cv.error)

# get MSE and compute RMSE
sqrt(ames_gbm1$cv.error[min_MSE])

gbm.perf(ames_gbm1, method = "cv")
```

# Tuning strategy 

Now let's tune the tree-specific hyperparameters

* we could do it in `caret` but lets use functional programming
* assess 3 values for tree depth
* assess 3 values for min obs in terminal node

___This grid search takes ~30 mins; remember, I said the ML process is more of a marathon than a sprint!!___

```{r slide-29}
# search grid
hyper_grid <- expand.grid(
  n.trees = 6000,
  shrinkage = .01,
  interaction.depth = c(3, 5, 7), #<<
  n.minobsinnode = c(5, 10, 15) #<<
)

model_fit <- function(n.trees, shrinkage, interaction.depth, n.minobsinnode) {
  set.seed(123)
  m <- gbm(
    formula = Sale_Price ~ .,
    data = ames_train,
    distribution = "gaussian",
    n.trees = n.trees,
    shrinkage = shrinkage, #<<
    interaction.depth = interaction.depth, #<<
    n.minobsinnode = n.minobsinnode,
    cv.folds = 10
  )
  # compute RMSE
  sqrt(min(m$cv.error))
}

hyper_grid$rmse <- pmap_dbl(
  hyper_grid,
  ~ model_fit(
    n.trees = ..1,
    shrinkage = ..2,
    interaction.depth = ..3,
    n.minobsinnode = ..4
    )
)

arrange(hyper_grid, rmse)
```

# Applying a Stochastic GBM

- start by assessing if values between 0.5-0.8 outperform your previous best model
- zoom in with a second round of tuning
- smaller values will tell you that overfitting was occurring

```{r slide-32}
bag_frac <- c(.5, .65, .8) #<<

for(i in bag_frac) {
  set.seed(123)
  m <- gbm(
    formula = Sale_Price ~ .,
    data = ames_train,
    distribution = "gaussian",
    n.trees = 6000, 
    shrinkage = 0.01, 
    interaction.depth = 7, 
    n.minobsinnode = 5,
    bag.fraction = i, #<<
    cv.folds = 10 
    )
  # compute RMSE
  print(sqrt(min(m$cv.error)))
}
```


# Extreme Gradient Boosting

# Prereqs

* __xgboost__ requires that our features are one-hot encoded
* __caret__ and __h2o::h2o.xgboost__ can automate this for you
* In this preprocessing I:
   - collapse low frequency levels to "other"
   - convert ordered factors to integers (aka label encode)
   
___Pro tip: If you have I cardinality categorical features, label or ordinal encoding often improves performance and speed!___

```{r slide-35}
library(recipes)
xgb_prep <- recipe(Sale_Price ~ ., data = ames_train) %>%
  step_other(all_nominal(), threshold = .005) %>%
  step_integer(all_nominal()) %>%
  prep(training = ames_train, retain = TRUE) %>%
  juice()

X <- as.matrix(xgb_prep[setdiff(names(xgb_prep), "Sale_Price")])
Y <- xgb_prep$Sale_Price
```

# First XGBoost model 

* `nrounds`: 6,000 trees 
* `objective`: `reg:linear` for regression but other options exist (i.e. `reg:logistic`, `binary:logistic`, `num_class`)
* `early_stopping_rounds`: stop training if CV RMSE doesn't improve for 50 trees in a row 
* `nfold`: 10-fold CV 

___This grid search takes ~20 secs___

```{r slide-36}
set.seed(123)
ames_xgb <- xgb.cv(
  data = X,
  label = Y,
  nrounds = 5000,
  objective = "reg:linear",
  early_stopping_rounds = 50, 
  nfold = 10,
  verbose = 0,
  )  

ames_xgb$evaluation_log %>% tail()
```

# Tuning 

1. ___Crank up the trees and tune learning rate with early stopping___
   - initial test RMSE results:
   - .red[`eta = .3` (default): 24,382 w/200 trees (< 1 min)]
   - .red[`eta = .1`: 22,333 w/398 trees (< 1 min)]
   - .green[`eta = .05`: 21,877 w/978 trees (1.5 min)]
   - .red[`eta = .01`: 22,094 w/2843 trees (4 min)]
   
As a comparison, if you one-hot encoded the feature set it takes 30 mins to run with `eta = .01`!

___This grid search takes ~1.5 min___

```{r slide-37}
set.seed(123)
ames_xgb <- xgb.cv(
  data = X,
  label = Y,
  nrounds = 6000,
  objective = "reg:linear",
  early_stopping_rounds = 50, 
  nfold = 10,
  verbose = 0,
  params = list(eta = .05) #<<
  )  

ames_xgb$evaluation_log %>% tail()
```

# Tuning 

1. Crank up the trees and tune learning rate with early stopping
2. ___Tune tree-specific hyperparameters___
   - tree depth
   - instances required to make additional split

* Preferred values: 
   - `max_depth` = 3
   - `min_child_weight` = 3
   - RMSE = 20989.27

___This grid search takes ~30 min___

```{r slide-38}
# grid
hyper_grid <- expand.grid(
  eta = .05,
  max_depth = c(1, 3, 5, 7, 9), #<<
  min_child_weight = c(1, 3, 5, 7, 9), #<<
  rmse = 0 # a place to dump results
  )

# grid search
for(i in seq_len(nrow(hyper_grid))) {
  set.seed(123)
  m <- xgb.cv(
    data = X,
    label = Y,
    nrounds = 6000,
    objective = "reg:linear",
    early_stopping_rounds = 50, 
    nfold = 10,
    verbose = 0,
    params = list( #<<
      eta = hyper_grid$eta[i], #<<
      max_depth = hyper_grid$max_depth[i], #<<
      min_child_weight = hyper_grid$min_child_weight[i] #<<
    ) #<<
  )
  hyper_grid$rmse[i] <- min(m$evaluation_log$test_rmse_mean)
}

arrange(hyper_grid, rmse)
```

# Tuning 

1. Crank up the trees and tune learning rate with early stopping
2. Tune tree-specific hyperparameters
3. ___Add stochastic attributes with___
   - subsampling rows for each tree
   - subsampling columns for each tree 

* Preferred values: 
   - `subsample` = 0.80
   - `colsample_bytree` = 1
   - RMSE = 20732.22

___This grid search takes ~12 min___

```{r slide-39}
# grid
hyper_grid <- expand.grid(
  eta = .05,
  max_depth = 3, 
  min_child_weight = 3,
  subsample = c(.5, .65, .8, 1), #<<
  colsample_bytree = c(.5, .65, .8, 1), #<<
  rmse = 0 # a place to dump results
  )

# grid search
for(i in seq_len(nrow(hyper_grid))) {
  set.seed(123)
  m <- xgb.cv(
    data = X,
    label = Y,
    nrounds = 6000,
    objective = "reg:linear",
    early_stopping_rounds = 50, 
    nfold = 10,
    verbose = 0,
    params = list( #<<
      eta = hyper_grid$eta[i],
      max_depth = hyper_grid$max_depth[i],
      min_child_weight = hyper_grid$min_child_weight[i],
      subsample = hyper_grid$subsample[i], #<<
      colsample_bytree = hyper_grid$colsample_bytree[i] #<<
    ) #<<
  )
  hyper_grid$rmse[i] <- min(m$evaluation_log$test_rmse_mean)
}

arrange(hyper_grid, rmse)
```


# Tuning

1. Crank up the trees and tune learning rate with early stopping
2. Tune tree-specific hyperparameters
3. Add stochastic attributes with
4. ___See if adding regularization helps___
   - gamma: tested 1, 100, 1000, 10000 -- no effect
   - lambda: tested 0.001, 0.01, 0.1, 1, 100, 1000 -- no effect
   - alpha: tested 0.001, 0.01, 0.1, 1, 100, 1000 -- minor improvement

* Preferred value:
   - `alpha` = 100
   - RMSE = 20581.31

___This grid search takes ~5 min___

```{r slide-41}
hyper_grid <- expand.grid(
  eta = .05,
  max_depth = 3, 
  min_child_weight = 3,
  subsample = .8, 
  colsample_bytree = 1,
  #gamma = c(1, 100, 1000, 10000),
  #lambda = c(1e-2, 0.1, 1, 100, 1000, 10000),
  alpha = c(1e-2, 0.1, 1, 100, 1000, 10000), #<<
  rmse = 0 # a place to dump results
  )

# grid search
for(i in seq_len(nrow(hyper_grid))) {
  set.seed(123)
  m <- xgb.cv(
    data = X,
    label = Y,
    nrounds = 6000,
    objective = "reg:linear",
    early_stopping_rounds = 50, 
    nfold = 10,
    verbose = 0,
    params = list( 
      eta = hyper_grid$eta[i], 
      max_depth = hyper_grid$max_depth[i],
      min_child_weight = hyper_grid$min_child_weight[i],
      subsample = hyper_grid$subsample[i], #<<
      colsample_bytree = hyper_grid$colsample_bytree[i],
      #gamma = hyper_grid$gamma[i], 
      #lambda = hyper_grid$lambda[i]#, 
      alpha = hyper_grid$alpha[i] #<<
    ) 
  )
  hyper_grid$rmse[i] <- min(m$evaluation_log$test_rmse_mean)
}

arrange(hyper_grid, rmse)
```

# Tuning 

1. Crank up the trees and tune learning rate with early stopping
2. Tune tree-specific hyperparameters
3. Add stochastic attributes with
4. See if adding regularization helps
5. If you find hyperparameter values that are substantially different from default settings, be sure to assess the learning rate again
6. ___Rerun final "optimal" model with `xgb.cv()` to get iterations required and then with `xgboost()` to produce final model___

___`final_cv`] test RMSE: 20,581.31___

```{r slide-42}
# parameter list
params <- list(
  eta = .05,
  max_depth = 3, 
  min_child_weight = 3,
  subsample = .8, 
  colsample_bytree = 1,
  alpha = 100
)

# final cv fit
set.seed(123)
final_cv <- xgb.cv(
  data = X,
  label = Y,
  nrounds = 6000,
  objective = "reg:linear",
  early_stopping_rounds = 50, 
  nfold = 10,
  verbose = 0,
  params = params #<<
  ) 

# train final model
ames_final_xgb <- xgboost(
  data = X,
  label = Y,
  nrounds = final_cv$best_iteration, #<<
  objective = "reg:linear",
  params = params, #<<
  verbose = 0
)
```

# Feature Importance

```{r slide-44}
vip::vip(ames_final_xgb, num_features = 25)
```

# Feature Effects (PDP)

```{r slide-45a}
ames_final_xgb %>%
  partial(
    pred.var = "Gr_Liv_Area", 
    n.trees = ames_final_xgb$niter, 
    grid.resolution = 50, 
    train = X
    ) %>%
  autoplot(rug = TRUE, train = X)
```

# Feature Effects (ICE)

```{r slide-45b}
ames_final_xgb %>%
  partial(
    pred.var = "Gr_Liv_Area", 
    n.trees = ames_final_xgb$niter, 
    grid.resolution = 50, 
    train = X,
    ice = TRUE
    ) %>%
  autoplot(rug = TRUE, train = X, alpha = .05, center = TRUE) 
```






